SoK: Web Application Attack Defenses (in Question)

### 2014-10-07 docs/questioning.tex
added attack section

\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{fancyhdr}
%\usepackage{lastpage}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{color}
\usepackage{url}
%\usepackage{moreverb}
\usepackage{verbatim}
\usepackage{textcomp}
\usepackage{mathptmx}
\usepackage{dingbat}
\usepackage{pifont}
\usepackage{acronym}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{supertabular}
\usepackage{listings}
\usepackage{threeparttable}
\usepackage{pdflscape}
\usepackage{array}
\usepackage{multirow}
\usepackage{subfigure}

@@

\title{SoK: Application Attack Defenses (in Question)}

@@

>>>>

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fancyhdr}
%\usepackage{lastpage}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{color}
\usepackage{url}
%\usepackage{moreverb}
\usepackage{verbatim}
\usepackage{textcomp}
\usepackage{mathptmx}
\usepackage{dingbat}
\usepackage{pifont}
\usepackage{acronym}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{supertabular}
\usepackage{listings}
\usepackage{threeparttable}
\usepackage{pdflscape}
\usepackage{array}
\usepackage{multirow}
\usepackage{subfigure}

@@

\title{SoK: Web Application Attack Defenses (in Question)}

@@

\section{Attacks}
\label{sec:attacks}

The basic problem behind web application attacks involves
the absence of input validation. By taking advantage of
this, attackers can inject either code into
applications to perform malicious tasks. Such
an exploit can have different forms depending on the
execution context of the application and the location
of the programming flaw that leads to the attack.

Bratus et al.~\cite{BLSPS11} portray the issue in a generic fashion:
{\it ``unexpected (and unexpectedly powerful) computational models
inside targeted systems, which turn a part of the target into a
so-called ``weird machine" programmable by the attacker via crafted inputs
(a.k.a. ``exploits")"}.
An example of the above definition is the
following: The code fragment below, defines the operation of
addition in the Scheme programming language~\cite{AS96,D09}:

\begin{verbatim}
(define (add x y) (+ x y))
\end{verbatim}

\noindent
Consider the case where instead of a number, a function that leads to an
endless loop is passed as an argument by the user. This will cause
the interpreter to enter
an endless loop and lead to a denial of service.
% Keep in mind that a {\sc cia} does not require a programming
% language with first-class functions.
The intuition here is that {\it ``every application that copies
untrusted input verbatim into an output
program is vulnerable to code injection"}.
Ray and Ligatti~\cite{RL12b} actually proved the above
claim based on formal language theory.

\begin{figure}
\begin{center}
\leavevmode
\includegraphics[scale=0.37]{attack-tree-uml.pdf}
\end{center}
\caption{\label{fig:taxonomy}A taxonomy of code injection attacks.
In our research we have excluded the defenses that deal with
the subcategories that are illustrated in grey color.}
\end{figure}

Figure \ref{fig:taxonomy} presents a taxonomy of code injection attacks divided
into two categories. The first involves binary code and the second
executable source code.

\subsection{Binary Code Injection}

Binary code injection attacks involve
the insertion of binary code into an application
to alter its execution flow and execute malicious
compiled code. This category involves buffer-overflow
attacks~\cite{CPMHWBBGWZ98,K11}, a staple of security problems.
An extensive survey on binary code
injection attacks can be found in reference~\cite{LC03}.
Furthermore, specific advances in exploiting such vulnerabilities
(i.e. ``heap smashing", ``arc injection" and others)
have been presented in reference~\cite{PB04}.
Finally, the countermeasures used to detect
such defects have already been surveyed~\cite{YJP12}
(many of them are also included in a book:~\cite{DKZ12}
-- Section 13.8), thus we do not include them in our work.

\subsection{Executable Source Code Injection}

Code injection also includes the use of source code, either of a
Domain Specific Language ({\sc dsl})
or a Dynamic Language. Note that binary code injection
attacks can only occur when the target system is implemented in
languages lacking array bounds checking, like
{\sc c} and {\sc c}++. Contrariwise, source code-driven
injection attacks can target applications written in various
programming languages with different features and characteristics.

Application attacks that involve {\sc dsl}s,
constitute an important subset of the code injection problem,
as {\sc dsl}s like {\sc sql} and {\sc xml} play an
significant role in the development of either web or mobile
applications. For example, many applications have interfaces
where a user enters input to interact with the
application's data, thereby interacting with the
underlying {\it Relational Database Management System}
({\sc rdbms}).
This input can become part of an
{\sc sql} query and executed on the target {\sc rdbms}.
A code injection attack that exploits the vulnerabilities of these
interfaces by taking advantage of input validation issues like
incorrectly passed parameters or incorrect type handling,
is called an ``{\sc sql} injection
attack''~\cite{CERT02,MS09,HVO06,SW06}.
By using similar techniques malicious users
can perform other exploits based on {\sc dsl}s,
like {\sc xp}ath~\cite{SW06,CDL07,MKS09} and
{\sc xml}~\cite{MSM13} injection attacks.

A recent class of {\sc cia}s involve
dynamic languages like Python, Perl, JavaScript, and {\sc php}~\cite{SFVM09,EWKK09,SMS13}.
In particular, JavaScript injection attacks comprise a wide subset of dynamic
language-driven attacks. Such attacks are manifested when a web application
accepts and redisplays data of uncertain origin without appropriate validation
and filtering. Based on this flaw, an attacker can
manage to inject a script in the JavaScript engine of a browser and alter its
execution flow~\cite{ELX07}. JavaScript injection attacks are considered as a
critical issue in web application security mainly because they are
associated with major vulnerabilities such as:
{\sc xss} attacks~\cite{SG07},
{\sc csrf} attacks~\cite{LZRL09}
and ``Cross-Channel Scripting"
({\sc xcs}) attacks~\cite{W10,BBB09}.

\section{Defenses for Detecting Web Application Attacks}

There are four main defensive approaches to counter
web application attacks, namely: {\it tainting},
{\it instruction set randomization} ({{\sc isr}}),
{\it policy enforcement} and {\it training}.
